---
title: "binomialDistribution"
output: html_document
date: "2023-10-03"
---

####Hyper geometry comes into picture if we have finite population like people,animals or products where we can identify the sample space being generated without replacement clearly.

#### All d function calculates the probability mass function and the p function calculates the cumulative function
#### An oil company has a 20% chance of striking oil when drilling a well. What is the probability the company strikes oil 5 out of the next 10 attempts
```{r}
n=10
p=0.2
s=5
dbinom(s,n,p)#Binomial distribution
```

#### An oil company has a 20% chance of striking oil when drilling a well. What is the probability the company strikes oil atmost 5 out of the next 10 attempts
p(X=0)+p(X=1)+p(X=2)+p(X=3)+p(X=4)+p(X=5)
```{r}
n=10
p=0.2
sum(dbinom(0:5,n,p))
dbinom(0,n,p)+dbinom(1,n,p)+dbinom(2,n,p)+dbinom(3,n,p)+dbinom(4,n,p)+dbinom(5,n,p)
pbinom(5,n,p)#Cumulative probability from X=0 to X=5
```
#### People arrive to donate blood. 42% of them have O type blood. What is the probability 30 of 120 have type O blood.
Hyper Geometry random variable

```{r}
dbinom(30,120,0.42)#Binomial Distribution

nsuccess=420000
nfailures=580000
n=120
j=30
dhyper(j,nsuccess,nfailures,n)#Hyper Geometry random variable
```
#### When sample size is too small when compared to the population then it's safe to switch between binomial distribution or hyper geometry random variable

```{r}
dbinom(3,4,0.99)
dhyper(3,9900,100,4)
```

```{r}
dbinom(4,9000,0.4)
dhyper(4,4000,6000,9000)
```
#### An oil drilling company has a 20% chance of striking oil when drilling a well. What is the probability that the company drills 7 times to strike oil the 3rd trial?
#### Negative Binomial Variable

```{r}
p=0.2
r=3
j=7
#For negative binomial we will not directly give the number of success trials we are interested in. Instead we will feed  the number of failures that we obtain before getting r successes.
dnbinom(j-r,r,p)
```
####In an airport, it is known that approximately 2 our of 10 passengers have a metallic object. If left undetected at manual security check at the airport entrance. Such a metallic object will raise an alarm when the passenger walks through the automated screening machine. It is considered as security breach if the alarm is raised for 20times a day. What is the probability of a security breach on a particular day when the total 100th passenger walks through the automated screening machine? 

```{r}
#Negative binomial
p=0.2
r=20
j=100
dnbinom(j-r,r,p)
```

#### There are 48 out of 50 itsems are defective. What is the probability of getting atleast 3 items defective when 45 itsems are taken.
```{r}
#Hyper geometry distribution
p=48/50
nsuccess=48
nfailure=2
n=45
#success X>=1 or 1-X=0
x=0
1-dhyper(x,nsuccess,nfailure,n)
```
#### A person conducting a telephone survey must get 4 or more surveys. But there is 90% chance that call gets rejected. What is the probability person has to do 10 more calls?

```{r}
#Negative binomial
p=0.10 #success rate = Call being accepted. 100-90=10%
j=10
r=4
dnbinom(j-r,r,p)
```
#### A certain stoplight, when coming from the North is green approximately 31% of the time. Over the next few days, someone comes to the light 8 times from North. We are interetsed in the probability that the person will come across green light 5 times.
```{r}
n=8
p=0.31
r=5
dbinom(r,n,p)
```
```{r}
n=8
p=0.31
sum(dbinom((5:8),n,p)) #P(x>=5)
pbinom(4,n,p) #P(X<=4)
1-pbinom(4,n,p) #P(X>=5)
```
#### An oil drilling company has a record of 2000 times of striking oil in 10000 times. What is the probability that the company strikes oil 7 times out of 20?

```{r}
n=20
nsuccess=2000
nfailures=8000
x=7
dhyper(x,nsuccess,nfailures,n)
```

#### An oil drilling company has a record of 4 times of striking oil in 10 times. What is the probability that the company strikes oil exactly 2 times, atleast 2 times and atmost 2 times.
```{r}
p=4/10
n=4
r=2
dbinom(r,n,p)#Exactly 2
pbinom(2,n,p)#Atmost 2
1-pbinom(1,n,p)#Atleast 2
```
#### rbinom is used to get it in the form of simulations.

```{r}
nsimulations=100000
simulatedData=rbinom(100000,n,p)
```

```{r}
simulatedData
```
#### Getting oil exactly 2 times
```{r}
mean(simulatedData==2)
```

#### Getting oil atleast 2 times
```{r}
mean(simulatedData>=2)
```
#### Getting oil atmost 2 times
```{r}
mean(simulatedData<=2)
```
#### Converting the simulatedData to a dataframe because the ggplot can work only on a table/data frame
```{r}
dfOil=as.data.frame(simulatedData)
colnames(dfOil)=c("OilDrilling")
head(dfOil)
```
```{r}
library(ggplot2)
```

#### Getting the plot for the frequencies
```{r}
p=ggplot(data=dfOil)+
  geom_bar(aes(x=OilDrilling),fill='steelblue',width=0.1)
```

```{r}
p
```
#### Converting the simulatedData to table so that we can get the frequencies as one column

```{r}
dfOil=as.data.frame(table(simulatedData))
head(dfOil)
```
#### Renaming the column names
```{r}
colnames(dfOil)=c("OilDrilling","Frequency")
head(dfOil)
```
#### Getting one more column with probability values
```{r}
dfOil["Probability"]=dfOil["Frequency"]/nsimulations
head(dfOil)

```
#### PMF - Probability mass function

```{r}
p=ggplot(data=dfOil)+
  geom_col(aes(x=OilDrilling,y=Probability),fill="steelblue",width=0.1)
```

```{r}
p
```

####Adding one more column with cumulative probability sum
```{r}
dfOil["cdf"]=cumsum(dfOil["Probability"])
head(dfOil)
```
##### Getting the plot for CDF - Cumulative Distribution Function using cumsum. P(x=0),P(X<=1),P(X<=2)....
```{r}
p=ggplot(data=dfOil)+
  geom_col(aes(x=OilDrilling,y=cdf),fill="red",width=0.1)
p
```



####Getting the plot for CDF - Cumulative Distribution Function. P(x=0),P(X<=1),P(X<=2)....
```{r}
n=4
p=0.4
cdf=pbinom(c(0:4),n,p)
p=ggplot(data=dfOil)+
  geom_col(aes(x=OilDrilling,y=cdf),fill="blue",width=0.1)
```

```{r}
p
```
```{r}
n=4
p=0.4
nsimulations=100
simulatedData=rsbinom(nsimulations,n,p)
mean(simulatedData)
```
```{r}
j=c(0:4)
k=dbinom(j,n,p)
k
sum(j*k)
```
#### Expected value of a binomial random variable to be n*p
#### Expected value of a hyper geometry random variable to be n*p
#### Expected value of a negative random variable to be r/p


