---
title: "Discrete Random Variables"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
```

---

**Scenario**: A salesman has scheduled two appointments to sell software, one in the morning and another one in the afternoon. There are two software editions available: the base edition costing Rs. 500 and the premium edition costing Rs. 1000. His morning appointments typically lead to a sale with a 30% chance while the afternoon ones typically lead to a sale with a 60% chance independent of what happened in the morning. If the morning appointment ends up in sale, the salesman has a 70% chance of selling the premium edition and if the afternoon appointment ends up in a sale, he is equally likely to sell either of the editions. Let X be the random variable representing the total Rupee value of sales. What are the different values that X can take? Calculate the probability that X takes the value 5000?

---


```{r}
# Sampling space for appointment success (0 corresponds to no sale, 1 corresponds to a sale)
s_appointment = ?

# Appointment success and failure probabilities
p_morning = ? # Success probability of sales in morning
p_afternoon =  ? # Success probability of sales in afternoon
p_appointment = matrix(nrow = 2, ncol = 2, c(1-p_morning, ?, ?, ?), byrow = TRUE)

# Sampling space for software type
s_software = c(?, ?)

# Software type probabilities
p_software = matrix(?, ?, ?, byrow = TRUE)

# Function that simulates one trial of the random experiment which is
# what the salesman earns on a random day
salesResult = function(){
  result = numeric(2)
  # Simulate whether sales happen in morning and afternoon appointments
  for (j in c(1:2)){
    result[j] = sample(?, size = 1, replace = TRUE, prob = p_appointment[?, ?])
  }
  earnings = ifelse(result[1] == 1, sample(?, size = 1, replace = TRUE, prob = p_software[?, ?]), 0) + ifelse(result[2] == 1, ?, ?)
  return(?)
}

# Number of simulations
nsimulations = 1e5
simulatedData = replicate(nsimulations, ?)

# Probability that the salesman earns Rs 5000
mean(simulatedData == ?)

# Can we calculate all the probabilities in one shot?
table(simulatedData) / ?

```

