---
title: "Graphing with ggplot"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Read Data

```{r}
foodData=read.csv('Data/food-texture.csv',header=TRUE,row.names=1)
head(foodData)
print(typeof(foodData))
print(class(foodData))
print(str(foodData))
```

#### Attributes of the Dataframe

```{r}
nrow(foodData)
ncol(foodData)
rownames(foodData)
colnames(foodData)
```

#### Accessing elements of the data

```{r}
foodData$Oil
foodData[['Oil']]
foodData[,'Oil']
foodData['Oil']
```

#### Factor data structure in R

```{r}
gender=c('Male','Female','Male','Female','Female','Male')
print(gender)
gender_categorical=factor(gender)
print(gender_categorical)
str(gender_categorical)
```

#### Adding new column

```{r}
#foodData['taste']=sample(c('poor','fair','good','excellent'),nrow(foodData),replace=TRUE,p=c(0.25,0.25,0.25,0.25))
#head(foodData)
```

#### write the updated data frame to food-texture.csv

```{r}
#write.csv(foodData,'Data/food-texture.csv')
```

#### Create a list of continuous and categorical feature names

```{r}
continuous_cols=c('Oil','Density','Fracture','Hardness')
categorical_cols=c('Crispy','taste')
```

#### Convert all categorical columns into factor type in one shot

```{r}
foodData[categorical_cols]=lapply((foodData[categorical_cols]),factor)
str(foodData)
```

#### Create a GGPlot canvas

```{r}
p=ggplot(data=foodData)

```

```{r}
library(ggplot2)
```

#### Create a visualization. Scatter plot

```{r}
p=ggplot(data=foodData) + 
  geom_point(aes(x=Density,y=Oil))
p
```

```{r}
p=ggplot(data=foodData) + 
  geom_point(aes(x=Density,y=Oil,color=Crispy))
p
```

```{r}
p=ggplot(data=foodData) + 
  geom_point(aes(x=Density,y=Oil,size=Crispy, color=Crispy))
p
```

```{r}
p=ggplot(data=foodData) + 
  geom_point(aes(x=Density,y=Oil,shape=taste, color=Crispy))
p
```

#### Visualization using box plot

```{r}
p=ggplot(data=foodData) + 
  geom_boxplot(aes(x=Crispy,y=Density))
p
```



```{r}
#foodData$Density
#foodData$Crispy==10
foodData[foodData$Crispy==10, ]
foodData[foodData$Crispy==10 & foodData$Oil<=17,]
foodData[foodData$Crispy==10, c('Density','Hardness')]
density10=foodData[foodData$Crispy==10,'Density']
density11=foodData[foodData$Crispy==11,'Density']
str(density10)
mean(density10)
#density10-mean(density10) #Broadcasting-mean centered vector
#(density10-mean(density10))^2 #Squared mean centered vector
#mean((density10-mean(density10))^2) #mean of the squared mean-centered vector-varience
#sqrt(mean((density10-mean(density10))^2))#Standard deviation of the vector

#z-score : z-score value tells how much above the value is from the average value

#var(density10) #varience
#sd(density10) #standard deviation
```

```{r}
density10=foodData[foodData$Crispy==10,'Density']
dfDensity10=data.frame(density10)
colnames(dfDensity10)=c('Density')
dfDensity10$DensityMC=density10-mean(density10)
dfDensity10$DensityMCS=(density10-mean(density10))^2




str(dfDensity10)
head(dfDensity10)
```


```{r}
#when plotting in ggplot it should be dataframe
dfDensity10=data.frame(density10)
colnames(dfDensity10)=c('Density')
str(dfDensity10)
```

```{r}
lapply(foodData[continuous_cols],scale)
```


####Component plot of density values of samples with crispy level is 10
```{r}
p=ggplot(data=dfDensity10)+
  geom_point(aes(x=c(1:nrow(dfDensity10)),y=Density),color='blue')+labs(x='Sample number',y='Density(kg/m^3)', title="Component plot of density")

```

```{r}
p=ggplot(data=dfDensity10)+
  geom_point(aes(x=c(1:nrow(dfDensity10)),y=DensityMC),color='Blue')+
geom_line(aes(x=c(1:nrow(dfDensity10)),y=mean(dfDensity10$DensityMC)),color='Red')+labs(x='Sample number',y='Density(kg/m^3)', title="Component plot of density")
p
```

```{r}
p=ggplot(data=dfDensity10)+
  geom_point(aes(x=c(1:nrow(dfDensity10)),y=DensityMC),color='Blue')+
geom_line(aes(x=c(1:nrow(dfDensity10)),y=mean(dfDensity10$DensityMC)),color='Red')+labs(x='Sample number',y='Density(kg/m^3)', title="Component plot of density")
p
```


```{r}
p=ggplot(data=dfDensity10)+
  geom_point(aes(x=c(1:nrow(dfDensity10)),y=Density),color='blue')+labs(x='Sample number',y='Density(kg/m^3)', title="Component plot of density")
```


#### Broadcasting - Subtracting mean of a value from all the values in a vector

```{r}
density10-mean(density10)
(density10-mean(density10))^2
mean((density10-mean(density10))^2) #Variance
sqrt(mean((density10-mean(density10))^2)) #Standard deviation
```

```{r}
# Load the mpg dataset
data('mpg')
mpgData = mpg

# Print the first five rows (or samples) in the data frame
head(mpgData, 5)
```

```{r}
# Initiate the ggplot() function binding to the car data frame
ggplot(data = mpgData)

# Create a plot object
p1 = ggplot(data = mpgData)

# Use the aes() function to specify the aesthetic mapping, that is, which variables
# should be plotted
p1 = ggplot(data = mpgData, aes(x = disp, y = hwy))

# Use the geom_ type functions to add geometric elements
p1 = ggplot(data = mpgData, aes(x = displ, y = hwy)) +
  geom_point()

# Add labels and title
p1 = p1 + labs(x = 'Displacement (litres)', y = 'Mileage (MPG)', title = 'Mileage vs. Displacement')
p1
```

```{r}
# Map aesthetics to variables

# Map the color aesthetic to the class variable
p2 = ggplot(data = mpgData) +
  geom_point(mapping = aes(x = displ, y = hwy))
p2
```

```{r}
# Map aesthetics to variables

# Map the size (or alpha or shape) aesthetic to the class variable
p3 = ggplot(data = mpgData) +
  geom_point(aes(x = displ, y = hwy))
p3
```

```{r}
# Set aesthetic manually

# Mark the points blue in color, square in shape, and fixed size
p4 = ggplot(data = mpgData) +
  geom_point(aes(x = displ, y = hwy))
p4
```

```{r}
# Add additional variables using facets

# Plot mileage w.r.t. each class individually
ggplot(data = mpgData) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(, nrow = 2)
```

```{r}
# Plot mileage w.r.t drive train and number of cylinders
ggplot(data = mpgData) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid()
```

```{r}
# A quick exercise on facet plotting with filtering of samples
# Investigate the number of levels in the cyl feature
factor()

# How many cars of each cyl type are there?
mpgData %>% 

# Filter samples with 3 and 5 cylinder cars (which are very rare)
# mpgData = mpgData %>% filter(!(? %in% c(?, ?)))
# 
# head(mpgData, 5)
# 
# # Map the color aesthetic to the cyl variable
# p5 = ggplot(data = mpgData) +
#   geom_point(mapping = aes(x = displ, y = hwy))
# p5
```
